---
title: "Daily Fluctuations in Teeth Brushing"
output: html_document
---

<style type="text/css">
h1 {
  font-size: 36px;
  margin-top: 60px;
  margin-bottom: 30px;
}

h2 {
  font-size: 28px;

}

h3 {
  font-size: 24px;
  padding-top: 30px;
  margin-bottom: 0px;
  text-align: left !important;
}

h4 {
  font-size: 18px;
}

p {
  font-size: 16px;
  text-align: left;
}

h1, h2, h3, h4, h5, h6 {
  text-align: center;
}

.p-text-align-right {
  text-align: right !important;
}

</style>

```{r, include=FALSE,warning=FALSE}
# clean workspace
rm(list=ls())

# load required packages
packages <- c("ggplot2", "ggpubr", "PerformanceAnalytics", "dplyr")
lapply(packages, library, character.only = TRUE)

# standardize ggplot2
theme_set(theme_bw() + theme( plot.title = element_text(size=18, lineheight=1, margin=margin(b=15, t=30), hjust = 0.5),
         axis.title.x = element_text(size=14, lineheight=1, margin=margin(t=15)),
         axis.title.y = element_text(size=14, lineheight=1, margin=margin(r=15))))

##################
## read in data ##
##################
daily_fluctuations<-read.csv("/Users/appollo_liu/Documents/workspace/Tooth_Brushing/data/clean_updated_brushing.csv", header=TRUE)

# valid_subjects <- c('119', '120', '121', '123', '124', '125', '127')
invalid_dates <- c('1/31/19', '2/1/19', '2/28/19', '3/1/19', '3/21/19', '3/22/19', '4/3/19', '4/4/19', '6/10/19', '6/11/19')
# take out all of the bad data
# uncomment if want to include specific subjects: `!(subject %in% valid_subjects)`
daily_fluctuations<-filter(daily_fluctuations, !(date %in% invalid_dates), serious_other_factors=="0", !is.na(time_brushing))

# makes each "key" in the data object its own variable
# Example: data$record_id is now record_id
attach(daily_fluctuations)

# find the overall participant information
# find the averages of daily fluctuations for each subject
average_time_brushing <- with(daily_fluctuations, ave(time_brushing, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_encouragements <- with(daily_fluctuations, ave(total_encouragements, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_parent_stress <- with(daily_fluctuations, ave(parent_stress, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_time_sleep <- with(daily_fluctuations, ave(time_sleep, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_parent_mood <- with(daily_fluctuations, ave(parent_mood, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_child_mood <- with(daily_fluctuations, ave(child_mood, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_time_sleep_nonap <- with(daily_fluctuations, ave(time_sleep_nonap, subject, FUN=function(x) mean(x, na.rm=TRUE)))

overall_subject <- data.frame("subject" = subject, "mean_time_brushing" = average_time_brushing, "mean_encouragements" = average_encouragements, "mean_parent_stress" = average_parent_stress, "mean_time_sleep" = average_time_sleep, "mean_parent_mood" = average_parent_mood, "mean_child_mood" = average_child_mood, "mean_time_sleep_nonap" = average_time_sleep_nonap)

overall_subject <- overall_subject[!duplicated(overall_subject$subject),]

# remove subject from the overall_subject data.frame because it is masked from daily_fluctuations
overall_subject <- overall_subject[,-1]
```

### Subject box plots

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject brushing length
brushing_length_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=time_brushing))
brushing_length_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Brushing Length", x="Subject",y="Brushing length (seconds)")
```

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject sleep
sleep_length_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=time_sleep))
sleep_length_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Sleep", x="Subject",y="Sleep (hours)")
```

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject's parent stress
parent_stress_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=parent_stress))
parent_stress_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Parent Stress", x="Subject",y="Parent stress level")
```

### Within Subject Relationships between Brushing length, Sleep, and Parent Stress

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot brushing length vs. sleep
within_subject_brushing_length_vs_sleep <- ggplot(data=daily_fluctuations, mapping=aes(x=time_sleep, y=time_brushing, color=as.factor(subject)))
within_subject_brushing_length_vs_sleep +
  stat_cor(method = "pearson",  size = 6, show.legend = FALSE) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Sleep", x="Sleep (hours)", y="Brushing length (seconds)", color="Subject") 
```

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot child sleep vs. parent stress
within_subject_child_sleep_vs_parent_stress <- ggplot(data=daily_fluctuations, mapping=aes(x=parent_stress, y=time_sleep, color=as.factor(subject)))
within_subject_child_sleep_vs_parent_stress +
  stat_cor(method = "pearson",  size = 6, show.legend = FALSE) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Child Sleep vs. Parent Stress", x="Parent stress", y=" Child sleep (hours)", color="Subject")
```

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot brushing length vs. parent stress
within_subject_brushing_length_vs_parent_stress <- ggplot(data=daily_fluctuations, mapping=aes(x=parent_stress, y=time_brushing, color=as.factor(subject)))
within_subject_brushing_length_vs_parent_stress +
  stat_cor(method = "pearson",  size = 6, show.legend = FALSE) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Parent Stress", x="Parent stress", y="Brushing length (seconds)", color="Subject")
```

### Between Subject Relationships between Brushing length, Sleep, Encouragement, and Parent Stress

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot encouragements vs. child sleep
between_subject_encouragements_vs_child_sleep <- ggplot(data=overall_subject, mapping=aes(x=mean_time_sleep, y=mean_encouragements))
between_subject_encouragements_vs_child_sleep +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Encouragement vs. Child Sleep", x="Child sleep (hours)", y="Encouragement")

# plot parent mood vs. parent stress
between_subject_parent_mood_vs_parent_stress <- ggplot(data=overall_subject, mapping=aes(x=mean_parent_stress, y=mean_parent_mood))
between_subject_parent_mood_vs_parent_stress +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Parent Mood vs. Parent Stress", x="Parent stress", y="Parent mood")

# plot brushing length vs. encouragements 
between_subject_brushing_length_vs_encouragements <- ggplot(data=overall_subject, mapping=aes(x=mean_encouragements, y=mean_time_brushing))
between_subject_brushing_length_vs_encouragements +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Encouragements", x="Encouragements", y="Brushing length")
```

### Daily Fluctuations Correlation Matrix

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
attach(overall_subject)
# cbind() combines two matrices that have the same number of rows together
combined_daily_fluctuations <- cbind(mean_time_brushing, mean_encouragements, mean_parent_stress, mean_parent_mood, mean_child_mood, mean_time_sleep, mean_time_sleep_nonap)
detach(overall_subject)

# graphing the correlation matrix
# pch controls the type of point that is displayed in the scatter plots
chart.Correlation(combined_daily_fluctuations, histogram=TRUE, pch=19)
```
### Possible Findings

<div style="height:60px;">
</div>