---
title: "Daily Fluctuations in Teeth Brushing"
output: html_document
---

<style type="text/css">
h1 {
  font-size: 36px;
  margin-top: 60px;
  margin-bottom: 30px;
}

h2 {
  font-size: 28px;

}

h3 {
  font-size: 24px;
  padding-top: 30px;
  margin-bottom: 0px;
  text-align: left !important;
}

h4 {
  font-size: 20px;
  padding-top: 30px;
}

p {
  font-size: 16px;
  text-align: left;
}

h1, h2, h3, h4, h5, h6 {
  text-align: center;
}

.p-text-align-right {
  text-align: right !important;
}

</style>

```{r, include=FALSE,warning=FALSE}
# clean workspace
rm(list=ls())

# load required packages
packages <- c("ggplot2", "ggpubr", "PerformanceAnalytics", "dplyr")
lapply(packages, library, character.only = TRUE)

# standardize ggplot2
theme_set(theme_bw() + theme( plot.title = element_text(size=18, lineheight=1, margin=margin(b=15, t=30), hjust = 0.5),
         axis.title.x = element_text(size=14, lineheight=1, margin=margin(t=15)),
         axis.title.y = element_text(size=14, lineheight=1, margin=margin(r=15))))

##################
## read in data ##
##################
subject_background_information<-read.csv("/Users/appollo_liu/Documents/workspace/Tooth_Brushing/data/parent_q_clean_n28.csv", header=TRUE)
daily_fluctuations<-read.csv("/Users/appollo_liu/Documents/workspace/Tooth_Brushing/data/clean_updated_brushing.csv", header=TRUE)

invalid_dates <- c('1/31/19', '2/28/19', '3/21/19', '4/3/19', '6/10/19')
# take out all of the data from the first day of the study, any data from an unusual day, and any data where there was no survey submitted for that day 
daily_fluctuations<-filter(daily_fluctuations, !(date %in% invalid_dates), other_stringent=="0", !is.na(parent_stress))

invalid_subjects <- c("103", "115", "126")
subject_background_information<-filter(subject_background_information, !(subject %in% invalid_subjects))


# makes each "key" in the data object its own variable
# Example: data$record_id is now record_id
attach(daily_fluctuations)

# find the overall participant information
overall_subject <- data.frame("subject_ID" = subject)
 
# find the averages of daily fluctuations for each subject
overall_subject$average_time_brushing <- with(daily_fluctuations, ave(time_brushing, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_time_brushing <- with(daily_fluctuations, ave(time_brushing, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_time_brushing <- overall_subject$sd_time_brushing/overall_subject$average_time_brushing

overall_subject$average_encouragements <- with(daily_fluctuations, ave(total_encouragements, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_encouragements <- with(daily_fluctuations, ave(total_encouragements, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_encouragements <- overall_subject$sd_encouragements/overall_subject$average_encouragements

overall_subject$average_parent_stress <- with(daily_fluctuations, ave(parent_stress, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_parent_stress <- with(daily_fluctuations, ave(parent_stress, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_parent_stress <- overall_subject$sd_parent_stress/overall_subject$average_parent_stress

overall_subject$average_time_sleep <- with(daily_fluctuations, ave(time_sleep, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_time_sleep <- with(daily_fluctuations, ave(time_sleep, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_time_sleep <- overall_subject$sd_time_sleep/overall_subject$average_time_sleep

overall_subject$average_parent_mood <- with(daily_fluctuations, ave(parent_mood, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_parent_mood <- with(daily_fluctuations, ave(parent_mood, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_parent_mood <- overall_subject$sd_parent_mood/overall_subject$average_parent_mood

overall_subject$average_child_mood <- with(daily_fluctuations, ave(child_mood, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_child_mood <- with(daily_fluctuations, ave(child_mood, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_child_mood <- overall_subject$sd_child_mood/overall_subject$average_child_mood

overall_subject$average_time_sleep_nonap <- with(daily_fluctuations, ave(time_sleep_nonap, subject, FUN=function(x) mean(x, na.rm=TRUE)))
overall_subject$sd_time_sleep_nonap <- with(daily_fluctuations, ave(time_sleep_nonap, subject, FUN=function(x) sd(x, na.rm=TRUE)))
overall_subject$cv_time_sleep_nonap <- overall_subject$sd_time_sleep_nonap/overall_subject$average_time_sleep_nonap

overall_subject <- overall_subject[!duplicated(overall_subject$subject_ID),]

# list of interesting data columns in subject_background_information
# age_firstword_recode_mos, parent1_edu, parent2_edu, income_median, num_siblings, child_birth_weight, age_walking_recode_mos
# average the parent education
overall_subject$average_parent_education <- (subject_background_information$parent1_edu + subject_background_information$parent2_edu) / 2

# subject 107 only provided parent education for parent 1, so making average education equal to parent 1 edu
overall_subject$average_parent_education <- replace(overall_subject$average_parent_education, 7, 18)

# add the subject_background information into the overall_subject data.frame
overall_subject$sum_pss <- subject_background_information$pss_sum
overall_subject$median_income <- subject_background_information$income_median
overall_subject$number_of_siblings <- subject_background_information$num_siblings
overall_subject$child_birth_weight <- subject_background_information$child_birth_weight
overall_subject$likes_brushing <- subject_background_information$like_brushing
```

### Subject box plots

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject brushing length
brushing_length_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=time_brushing))
brushing_length_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Brushing Length", x="Subject",y="Brushing length (seconds)")

# plot subject sleep
sleep_length_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=time_sleep))
sleep_length_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Sleep", x="Subject",y="Sleep (hours)")
```

### Within Subject Relationships

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot time brushing vs. encouragements
within_subject_time_brushing_vs_encouragements <- ggplot(data=daily_fluctuations, aes(x=total_encouragements, y=time_brushing))
within_subject_time_brushing_vs_encouragements + 
  facet_wrap(~subject) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE, fullrange=FALSE) +
  stat_cor(method = "pearson",  size =3) +
  labs(title="Time Brushing vs. Encouragements", x="Encouragements",y="Time brushing (seconds)")
```

### Within Subject Correlation Matrices

#### Correlation Matrix For Each Valid Subjects
```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# unique(daily_fluctuations$subject)
for (subject_id in unique(daily_fluctuations$subject)) {
  print(paste("Correlation matrix for subject =", toString(subject_id), sep=" "))
  subject_daily_fluctuations <- filter(daily_fluctuations, subject==subject_id)
  chart.Correlation(dplyr::select(subject_daily_fluctuations, time_brushing, total_encouragements, parent_stress, parent_mood, child_mood, time_sleep), histogram=TRUE, pch=19)
}
```

<p>
  Note: The symbols show these p-values. 0.001: 3 stars, 0.01: 2 stars, 0.05: 1 star, 0.10: ".")
</p>

#### Correlation Matrix For Subjects In Study Version 2
```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# unique(daily_fluctuations$subject)
for (subject_id in unique(daily_fluctuations$subject)) {
  if (as.numeric(subject_id) > 118) {
    # part of study version 2
    print(paste("Correlation matrix for subject =", toString(subject_id), sep=" "))
    subject_daily_fluctuations <- filter(daily_fluctuations, subject==subject_id)
    chart.Correlation(dplyr::select(subject_daily_fluctuations, time_brushing, total_encouragements, parent_stress, parent_mood, child_mood, time_sleep, fussy_am, fussy_pm_reversed), histogram=TRUE, pch=19)
  }
}
```

#### Looking at Child Brushing Extremes

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
worst_time_brushing<-filter(daily_fluctuations, brushing_rank=="worst")
best_time_brushing<-filter(daily_fluctuations, brushing_rank=="best")

between_subject_worst_brushing_sleep_vs_average_sleep <- ggplot()
between_subject_worst_brushing_sleep_vs_average_sleep +
  geom_point(data=overall_subject, aes(x=subject_ID, y=average_time_sleep), shape=5, size=4) +
  geom_point(data=worst_time_brushing, aes(x=subject, y=time_sleep), shape=1, size=4) +
  labs(title="Subject Worst Night Brushing Sleep vs. Average Sleep", x="Subject", y="Child sleep (hours)", color="Sleep")
```

```{r,warning=FALSE}
sum(overall_subject$average_time_sleep > worst_time_brushing$time_sleep)
length(overall_subject$average_time_sleep)
```

### Between Subject Relationships

#### Between Subject Correlation Matrices

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE, warn.conflicts = FALSE}
attach(overall_subject)
# cbind() combines two matrices that have the same number of rows together
between_subject_averages <- cbind(average_time_brushing, average_encouragements, average_parent_stress, average_parent_mood, average_child_mood, average_time_sleep, average_parent_education, median_income, number_of_siblings, child_birth_weight, sum_pss, likes_brushing)

between_subject_average_and_sd <- cbind(average_time_brushing, sd_time_brushing, average_encouragements, sd_encouragements, average_parent_stress, sd_parent_stress, average_parent_mood, sd_parent_mood, average_child_mood, sd_child_mood, average_time_sleep, sd_time_sleep)

between_subject_cv <- cbind(cv_time_brushing, cv_encouragements, cv_parent_stress, cv_time_sleep, cv_parent_mood, cv_child_mood)
detach(overall_subject)

# graphing the correlation matrix
# pch controls the type of point that is displayed in the scatter plots
chart.Correlation(between_subject_averages, histogram=TRUE, pch=19)

chart.Correlation(between_subject_average_and_sd, histogram=TRUE, pch=19)

chart.Correlation(between_subject_cv, histogram=TRUE, pch=19)
```
<p>
  Notable correlations in the first between subject correlation matrix:  
    1. Negative correlation between average time brushing and pss sum which supports the similar negative correlation between average time brushing and average parent stress.  
    2. There’s also an interesting positive correlation between number of siblings and child birth weight.  
    3. A positive correlation between time brushing and how much the child likes brushing.  
    4. Positive correlation between time brushing and median income.  
    5. Very large negative correlation between time brushing and number of siblings.  
Note: Correlations 3 through 5 only take into account responses from subjects in second version of study.
</p>
<p>
  List of all correlations in mean and standard deviation (2nd) correlation matrix:  
    1. average_time_brushing & sd_time_brushing: +  
    2. average_time_brushing & average_encouragement: +  
    3. average_time_brushing & sd_encouragement: +. Makes sense, the more child brushing, the more variation in how much time the parent has to give more encouragements.  
    4. average_time_brushing & average_parent_stress: -  
    5. sd_time_brushing & average_encouragement: +. The greater the variation in brushing, the more encouragement. This could be that because the parent knows that the child has the potential to brush more, the parent can try to offer more encouragement to the child so that the child is reaching their "highest potential" time brushing.  
    6. sd_time_brushing & sd_encouragement: +  
    7. sd_time_brushing & average_parent_stress: -. Higher variation in brushing correlates with lower average parent stress. Would have thought that higher variation would cause parent stress to be higher because the parent doesn't know what's going to happen.  
    8. average_encouragement & sd_encouragement: +. Similar to correlation to average_time_brushing & sd_time_brushing: the greater the average number of encouragements, the greater the variation in the total number of encouragements.  
    9. sd_encouragement & average_child_mood: -. Higher variation in encouragment correlates with lower average child mood. Kind of makes sense: if the child is unhappy, parents may react differently by either giving a lot of encouragement or give no encouragement and give up.  
    10. average_parent_stress & average_parent_mood: -  
    11. average_parent_stress & average_child_mood: -  
    12. sd_parent_stress & sd_parent_mood: +  
    13. average_parent_mood & sd_parent_mood: -  
    14. average_parent_mood & average_child_mood: +  
    15. average_parent_mood & sd_time_sleep: +. High average parent mood correlates with high variation in child sleep. Not really sure because would have assummed that low variation in child sleep would make the parent happier.  
    16. sd_parent_mood & sd_child_mood: +  
    17. sd_parent_mood & average_time_sleep: -. High variation in parent mood correlates with lower child sleep. Kind of makes sense: parent mood varies more when the child sleeps less.  
    18. average_child_mood & sd_child_mood: -. High average child mood correlates with less variation in the child mood.  
</p>
<p>
  Interesting finding that the more variation in the child's mood, the more variation there is in time brushing according to the cv correlation matrix.
</p>
<div style="height:60px;">
</div>