---
title: "Daily Fluctuations in Teeth Brushing"
output: html_document
---

<style type="text/css">
h1 {
  font-size: 36px;
  margin-top: 60px;
  margin-bottom: 30px;
}

h2 {
  font-size: 28px;

}

h3 {
  font-size: 24px;
  padding-top: 30px;
  margin-bottom: 0px;
  text-align: left !important;
}

h4 {
  font-size: 18px;
}

p {
  font-size: 16px;
  text-align: left;
}

h1, h2, h3, h4, h5, h6 {
  text-align: center;
}

.p-text-align-right {
  text-align: right !important;
}

</style>

```{r, include=FALSE,warning=FALSE}
# clean workspace
rm(list=ls())

# load required packages
packages <- c("ggplot2", "ggpubr", "PerformanceAnalytics", "dplyr")
lapply(packages, library, character.only = TRUE)

# standardize ggplot2
theme_set(theme_bw() + theme( plot.title = element_text(size=18, lineheight=1, margin=margin(b=15, t=30), hjust = 0.5),
         axis.title.x = element_text(size=14, lineheight=1, margin=margin(t=15)),
         axis.title.y = element_text(size=14, lineheight=1, margin=margin(r=15))))

##################
## read in data ##
##################
daily_fluctuations<-read.csv("/Users/appollo_liu/Documents/workspace/Tooth_Brushing/data/clean_updated_brushing.csv", header=TRUE)

# valid_subjects <- c('119', '120', '121', '123', '124', '125', '127')
invalid_dates <- c('1/31/19', '2/28/19', '3/21/19', '4/3/19', '6/10/19')
# take out all of the data from the first two days of the study, any data from an unusual day, and any data wehre there was no survey submitted for that day 
# uncomment if want to include specific subjects: `!(subject %in% valid_subjects)`
daily_fluctuations<-filter(daily_fluctuations, !(date %in% invalid_dates), other_stringent=="0", !is.na(parent_stress))

# makes each "key" in the data object its own variable
# Example: data$record_id is now record_id
attach(daily_fluctuations)

# find the overall participant information
# find the averages of daily fluctuations for each subject
average_time_brushing <- with(daily_fluctuations, ave(time_brushing, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_encouragements <- with(daily_fluctuations, ave(total_encouragements, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_parent_stress <- with(daily_fluctuations, ave(parent_stress, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_time_sleep <- with(daily_fluctuations, ave(time_sleep, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_parent_mood <- with(daily_fluctuations, ave(parent_mood, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_child_mood <- with(daily_fluctuations, ave(child_mood, subject, FUN=function(x) mean(x, na.rm=TRUE)))
average_time_sleep_nonap <- with(daily_fluctuations, ave(time_sleep_nonap, subject, FUN=function(x) mean(x, na.rm=TRUE)))

overall_subject <- data.frame("subject_ID" = subject, "mean_time_brushing" = average_time_brushing, "mean_encouragements" = average_encouragements, "mean_parent_stress" = average_parent_stress, "mean_time_sleep" = average_time_sleep, "mean_parent_mood" = average_parent_mood, "mean_child_mood" = average_child_mood, "mean_time_sleep_nonap" = average_time_sleep_nonap)

overall_subject <- overall_subject[!duplicated(overall_subject$subject_ID),]
```

### Subject box plots

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject brushing length
brushing_length_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=time_brushing))
brushing_length_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Brushing Length", x="Subject",y="Brushing length (seconds)")
```

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject sleep
sleep_length_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=time_sleep))
sleep_length_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Sleep", x="Subject",y="Sleep (hours)")
```

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject's parent stress
parent_stress_vs_subject <- ggplot(data=daily_fluctuations, mapping=aes(x=as.factor(subject), y=parent_stress))
parent_stress_vs_subject +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Parent Stress", x="Subject",y="Parent stress level")
```

### Within Subject Relationships between Brushing length, Sleep, and Parent Stress

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot time brushing vs. encouragements
within_subject_time_brushing_vs_encouragements <- ggplot(data=daily_fluctuations, aes(x=total_encouragements, y=time_brushing))
within_subject_time_brushing_vs_encouragements + 
  facet_wrap(~subject) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE, fullrange=FALSE) +
  stat_cor(method = "pearson",  size =3) +
  labs(title="Time Brushing vs. Encouragements", x="Encouragements",y="Time brushing (seconds)")

# plot time brushing vs. time sleep
within_subject_time_brushing_vs_time_sleep <- ggplot(data=daily_fluctuations, aes(x=time_sleep, y=time_brushing))
within_subject_time_brushing_vs_time_sleep + 
  facet_wrap(~subject) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE, fullrange=FALSE) +
  stat_cor(method = "pearson",  size =3) +
  labs(title="Time Brushing vs. Child Sleep", x="Child sleep (hours)",y="Time brushing (seconds)")

# plot time_brushing vs. child mood
within_subject_time_brushing_vs_child_mood <- ggplot(data=daily_fluctuations, aes(x=child_mood, y=time_brushing))
within_subject_time_brushing_vs_child_mood + 
  facet_wrap(~subject) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE, fullrange=FALSE) +
  stat_cor(method = "pearson",  size =3) +
  labs(title="Time Brushing vs. Child Mood", x="Child mood", y="Time brushing (seconds)")

# plot encouragements vs. child sleep
within_subject_encouragements_vs_child_sleep <- ggplot(data=daily_fluctuations, aes(x=time_sleep, y=total_encouragements))
within_subject_encouragements_vs_child_sleep + 
  facet_wrap(~subject) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE, fullrange=FALSE) +
  stat_cor(method = "pearson",  size =3) +
  labs(title="Encouragements vs. Child Sleep", x="Child sleep (hours)",y="Encouragements")

# plot encouragements vs. child mood
within_subject_encouragements_vs_child_mood <- ggplot(data=daily_fluctuations, aes(x=child_mood, y=total_encouragements))
within_subject_encouragements_vs_child_mood + 
  facet_wrap(~subject) +
  geom_point(shape=1) +
  geom_smooth(method=lm, se=FALSE, fullrange=FALSE) +
  stat_cor(method = "pearson",  size =3) +
  labs(title="Encouragements vs. Child Mood", x="Child mood",y="Encouragements")
```

### Between Subject Relationships

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# has unusually data
# overall_subject<-filter(overall_subject, subject!="120", subject!="101")

# plot brushing length vs. encouragements 
between_subject_brushing_length_vs_encouragements <- ggplot(data=overall_subject, mapping=aes(x=mean_encouragements, y=mean_time_brushing))
between_subject_brushing_length_vs_encouragements +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Encouragements", x="Encouragements", y="Brushing length")

# plot brushing length vs. child sleep
between_subject_brushing_length_vs_child_sleep <- ggplot(data=overall_subject, mapping=aes(x=mean_time_sleep, y=mean_time_brushing))
between_subject_brushing_length_vs_child_sleep +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Child Sleep", x="Child sleep (hours)", y="Brushing length (seconds)")

# plot brushing length vs. child mood
between_subject_brushing_length_vs_child_mood <- ggplot(data=overall_subject, mapping=aes(x=mean_child_mood, y=mean_time_brushing))
between_subject_brushing_length_vs_child_mood +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Child Mood", x="Child sleep mood", y="Brushing length (seconds)")

# plot encouragements vs. child sleep
between_subject_encouragements_vs_time_sleep <- ggplot(data=overall_subject, mapping=aes(x=mean_time_sleep, y=mean_encouragements))
between_subject_encouragements_vs_time_sleep +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Encouragement vs. Child sleep", x="Child sleep (hours)", y="Encouragement")

# plot encouragements vs. child mood
between_subject_encouragements_vs_child_mood <- ggplot(data=overall_subject, mapping=aes(x=mean_child_mood, y=mean_encouragements))
between_subject_encouragements_vs_child_mood +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Encouragement vs. Child Mood", x="Child mood", y="Encouragement")
```

### Daily Fluctuations Correlation Matrix

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
attach(overall_subject)
# cbind() combines two matrices that have the same number of rows together
combined_daily_fluctuations <- cbind(mean_time_brushing, mean_encouragements, mean_parent_stress, mean_parent_mood, mean_child_mood, mean_time_sleep, mean_time_sleep_nonap)
detach(overall_subject)

# graphing the correlation matrix
# pch controls the type of point that is displayed in the scatter plots
chart.Correlation(combined_daily_fluctuations, histogram=TRUE, pch=19)
```

### Looking at Child Brushing Extremes

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
worst_time_brushing<-filter(daily_fluctuations, brushing_rank=="worst")
best_time_brushing<-filter(daily_fluctuations, brushing_rank=="best")

between_subject_worst_brushing_sleep_vs_mean_sleep <- ggplot()
between_subject_worst_brushing_sleep_vs_mean_sleep +
  geom_point(data=overall_subject, aes(x=subject_ID, y=mean_time_sleep), shape=5, size=4) +
  geom_point(data=worst_time_brushing, aes(x=subject, y=time_sleep), shape=1, size=4) +
  labs(title="Subject Worst Night Brushing Sleep vs. Average Sleep", x="Subject", y="Child sleep (hours)", color="Sleep")
```

<div style="height:60px;">
</div>