---
title: "Daily Fluctuations in Teeth Brushing Pilot 4"
output: html_document
---

<style type="text/css">
h1 {
  font-size: 36px;
  margin-top: 60px;
  margin-bottom: 30px;
}

h2 {
  font-size: 28px;

}

h3 {
  font-size: 24px;
  padding-top: 30px;
  margin-bottom: 0px;
  text-align: left !important;
}

h4 {
  font-size: 16px;
}

h5 {
 font-size: 14px;
}

p {
  font-size: 12px;
  text-align: left;
  line-height: 1;
}

h1, h2, h3, h4, h5, h6 {
  text-align: center;
  line-height: 1;
}

.p-text-align-right {
  text-align: right !important;
}

</style>

```{r, include=FALSE,warning=FALSE}
# clean workspace
rm(list=ls())

# load required packages
packages <- c("ggplot2", "ggpubr", "PerformanceAnalytics")
lapply(packages, library, character.only = TRUE)

# standardize ggplot2
theme_set(theme_bw() + theme( plot.title = element_text(size=18, lineheight=1, margin=margin(b=15, t=30), hjust = 0.5),
         axis.title.x = element_text(size=14, lineheight=1, margin=margin(t=15)),
         axis.title.y = element_text(size=14, lineheight=1, margin=margin(r=15))))

##################
## read in data ##
##################
participant_background_information<-read.csv("/Users/appollo_liu/Documents/workspace/Tooth_Brushing/data/parent_q_clean_n19.csv", header=TRUE)
daily_fluctuations<-read.csv("/Users/appollo_liu/Documents/workspace/Tooth_Brushing/data/pilot_041919_merged_n19.csv", header=TRUE)

data<-merge(participant_background_information, daily_fluctuations, by="subject")
# exclude subject 115
data<-data[!grepl(115, data$subject),]

# makes each "key" in the data object its own variable
# Example: data$record_id is now record_id
attach(data)
```

### Subject box plots

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject brushing length
brushing_length <- ggplot(data=data, mapping=aes(x=as.factor(subject), y=time_brushing))
brushing_length +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Brushing Length", x="Subject",y="Brushing length (seconds)")
```  

<p>
  Some subjets seem to have large differences in brushing length between days (subjects 101, 102, 104, 106, 110, 113, 116, 118) which is shown by the large difference between the third and first quartile. Other subjects like 100, 112, 114, and 117 have smaller interquartile ranges, suggesting that their brushing lengths do not vary as much.
</p>

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject sleep
sleep_length <- ggplot(data=data, mapping=aes(x=as.factor(subject), y=time_sleep))
sleep_length +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Subject Sleep", x="Subject",y="Sleep (hours)")
```

<p>
  The majority of subjects also seem to vary in the amount of sleep they get, ranging from about 9 to 13 hours with the exception of subject 103 who had a relatively stable 11.5 hours of sleep with a few nights that were exceptions.
</p>

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot subject's parent stress
parent_stress <- ggplot(data=data, mapping=aes(x=as.factor(subject), y=parent_stress))
parent_stress +
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=9, size=3) +
  labs(title="Parent Stress", x="Subject",y="Parent stress level")
```

<p>
  Similar to child sleep length, parent stress level varies from day to day with the exception of subject 106's parent. Note: this only accounts for the stress level of the parent submitting the survey and not both parents.
</p>

### Relationships between Brushing length, Sleep, and Parent Stress

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# plot brushing length vs. sleep
brushing_length_vs_sleep <- ggplot(data=data, mapping=aes(x=time_sleep, y=time_brushing, color=as.factor(subject)))
brushing_length_vs_sleep +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Sleep", x="Sleep (hours)", y="Brushing length (seconds)", color="Subject")
  
# plot sleep vs. parent stress
sleep_vs_parent_stress <- ggplot(data=data, mapping=aes(x=parent_stress, y=time_sleep, color=as.factor(subject)))
sleep_vs_parent_stress +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Sleep vs. Parent Stress", x="Parent stress", y="Sleep (hours)", color="Subject")

# plot brushing length vs. parent stress
brushing_length_vs_parent_stress <- ggplot(data=data, mapping=aes(x=parent_stress, y=time_brushing, color=as.factor(subject)))
brushing_length_vs_parent_stress +
  stat_cor(method = "pearson",  size = 6) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  geom_point(shape=1) +
  labs(title="Brushing Length vs. Parent Stress", x="Parent stress", y="Brushing length (seconds)", color="Subject")
```

### Daily Fluctuations Correlation Matrix

```{r,warning=FALSE,fig.width=10, fig.height=8,echo = FALSE}
# cbind() combines two matrices that have the same number of rows together
combined_daily_fluctuations<-cbind(time_brushing, encouragement, parent_stress, parent_mood, child_mood, time_sleep, time_sleep_nonap)

# graphing the correlation matrix
# pch controls the type of point that is displayed in the scatter plots
chart.Correlation(combined_daily_fluctuations, histogram=TRUE, pch=19)
```